Surface correlation functions calculations: improving methodology for discrete images with novel exact continuous approach

### 2023-03-24 paper.tex
Improve section about dual numbers

\subsection{Dual numbers and automatic differentiation}
\label{sec:dual}
A ring of dual numbers $D$ is a set of pairs of real numbers $(x, y)$ with two
binary operations: $+$ (addition) and $\cdot$ (multiplication) where
multiplication works according to the following law:
\begin{equation*}
  (x_1, y_1)\cdot(x_2, y_2) = (x_1x_2, x_1y_2 + y_1x_2)
\end{equation*}
It's easy to see that $D$ is commutative.

It's useful to introduce an imaginary unit $\varepsilon$ and write a dual number
$(x, y)$ as $x + y\varepsilon$. As you can see from multiplication law,
$1\cdot \varepsilon = \varepsilon$ and $\varepsilon^2 = 0$, hence $D$ is not an
integral domain and division is not defined \textcolor{red}{for all dual numbers with exception
of $0$.} Namely it's defined only if divisor has non-zero real part:
\begin{align*}
  \frac{a+b\varepsilon}{c+d\varepsilon} &=
  \frac{(a+b\varepsilon)(c-d\varepsilon)}{(c+d\varepsilon)(c-d\varepsilon)} \\
  &= \frac{ac-ad\varepsilon+bc\varepsilon}{c^2} \\
  &= \frac{a}{c} + \frac{bc-ad}{c^2}\varepsilon
\end{align*}

Power function $(a + b\varepsilon)^n$ where $n \in \mathbb{N}$ can be defined
using binomial formula and keeping in mind that $\varepsilon^n = 0$ if $n>1$:
\begin{equation*}
  (a + b\varepsilon)^n = \sum_{k=0}^n \binom{n}{k} a^k (b\varepsilon)^{n-k} =
  a^n + n a^{n-1} b \varepsilon
\end{equation*}
This formula can be extended on $n \in \mathbb{Q}$ as well.

Elementary functions can be extended to dual numbers using Taylor series. For
example, the exponent function is extended as:
\begin{align*}
  e^{a + b\varepsilon} &= \sum_{k=0}^{\infty}\frac{(a + b\varepsilon)^k}{k!} \\
  &= \sum_{k=0}^{\infty} \frac{a^k + bka^{k-1}\varepsilon}{k!} \\
  &= \sum_{k=0}^{\infty} \frac{a^k}{k!} + b\varepsilon \sum_{k=1}^{\infty}
  \frac{a^{k-1}}{(k-1)!} \\
  &= e^a + e^ab\varepsilon
\end{align*}

@@

It's easy to see, that there is a connection between dual numbers and
differentiation. Indeed, if functions $f(x)$ and $g(x)$
are differentiable at the point $a$ and $h(x) = f(x)g(x)$ is their product then
$h(a + b\varepsilon) = h(a) + b\varepsilon(f'(a)g(a) + f(a)g'(a))$.
If $h(x) = f(x)/g(x)$ and $g(a) \ne 0$ then
$h(a + b\varepsilon) = h(a) + b\varepsilon\frac{f'(a)g(a) - f(a)g'(a)}{f^2(a)}$.
More generally, we can see that $f(a + b\varepsilon) = f(a) + bf'(a)\varepsilon$.

So, using dual numbers we can compute two values at once: one of the function at
the point $x = a$ and the other of its derivative at this point:
$f(a + \varepsilon) = f(a) + \varepsilon f'(a)$. Also we can compute a partial
derivative of a multivariate function. For example:
\begin{align*}
  \frac{\partial}{\partial x} f(x, y) \vert_{x = a, y = b} &= f(a + \varepsilon,
  b) \\
  \frac{\partial}{\partial y} f(x, y) \vert_{x = a, y = b} &= f(a, b +
  \varepsilon)
\end{align*}

@@

Summarizing, the algorithm of calculating $F_{ss}(a, b)$ is the following:
\begin{enumerate}
\item Find $X$ which is a set of soulutions of \cref{eq:inter}.
\item For all points in $X$ find normals to the boundaries using automatic
  differentiation.
\item Calculate $F_{ss}(a, b)$ using \cref{eq:fss-2d}.
\end{enumerate}

>>>>

\subsection{Dual numbers and automatic differentiation}
\label{sec:dual}
A ring of dual numbers $D$ is a set of pairs of real numbers $(a, b)$ with two
binary operations: $+$ (addition) and $\cdot$ (multiplication) where
multiplication works according to the following law:
\begin{equation*}
  (a, b)\cdot(c, d) = (ac, ad + bc)
\end{equation*}
It's easy to see that $D$ is commutative.

It's useful to introduce an imaginary unit $\varepsilon$ and write a dual number
$(a, b)$ as $a + b\varepsilon$. As you can see from multiplication law,
$1\cdot \varepsilon = \varepsilon$ and $\varepsilon^2 = 0$, hence $D$ is not an
integral domain and division is defined only if divisor has non-zero real part:
\begin{align*}
  \frac{a+b\varepsilon}{c+d\varepsilon} &=
  \frac{(a+b\varepsilon)(c-d\varepsilon)}{(c+d\varepsilon)(c-d\varepsilon)} \\
  &= \frac{ac-ad\varepsilon+bc\varepsilon}{c^2} \\
  &= \frac{a}{c} + \frac{bc-ad}{c^2}\varepsilon
\end{align*}

Now we consider functions of dual variable $x + y\varepsilon$. Power function
$(x + y\varepsilon)^n$ where $n \in \mathbb{N}$ can be defined using binomial
formula and keeping in mind that $\varepsilon^n = 0$ if $n>1$:
\begin{equation*}
  (x + y\varepsilon)^n = \sum_{k=0}^n \binom{n}{k} x^k (y\varepsilon)^{n-k} =
  x^n + n x^{n-1} y \varepsilon
\end{equation*}
This formula can be extended to $n \in \mathbb{Q}$ as well.

Elementary functions can be extended to dual numbers using Taylor series (here
$a$ is some point at which $f$ is differentiable):
\begin{align*}
  f(x + y\varepsilon) &= f(a) + f'(a)(x + y\varepsilon) + \frac{f''(a)}{2!}(x + y\varepsilon)^2 \\
  & \qquad + \frac{f'''(a)}{3!}(x + y\varepsilon)^3 + \dots \\
  &= f(a) + f'(a)(x + y\varepsilon) + \frac{f''(a)}{2!}(x^2 + 2xy\varepsilon) \\
  & \qquad + \frac{f'''(a)}{3!}(x^3 + 3x^2y\varepsilon) + \dots \\
  &= f(a) + f'(a)x + \frac{f''(a)}{2!}x^2 + \frac{f'''(a)}{3!}x^3 + \dots \\
  & \quad y\varepsilon(f'(a) + f''(a)x + \frac{f'''(a)}{2!}x^2 + \dots) \\
  &= f(x) + yf'(x)\varepsilon
\end{align*}

@@

As you can see, dual numbers are connected with differentiation. Using a dual
argument $a + \varepsilon$ we can compute a value of a function at point $x = a$
and its derivative at that point:
$f(a + \varepsilon) = f(a) + f'(a)\varepsilon$. Also we can compute a partial
derivative of a multivariate function. For example:
\begin{equation}
  \begin{aligned}
    \frac{\partial}{\partial x} f(x, y) \vert_{x = a, y = b} &= f(a + \varepsilon,
    b) \\
    \frac{\partial}{\partial y} f(x, y) \vert_{x = a, y = b} &= f(a, b +
    \varepsilon)
  \end{aligned}
  \label{eq:autonormals}
\end{equation}
The last equation will be useful for computation of normals to the interface in
next sections.

@@

Summarizing, the algorithm of calculating $F_{ss}(a, b)$ is the following:
\begin{enumerate}
\item Find $X$ which is a set of soulutions of \cref{eq:inter}.
\item For all points in $X$ find normals to the boundaries using
  \cref{eq:autonormals}.
\item Calculate $F_{ss}(a, b)$ using \cref{eq:fss-2d}.
\end{enumerate}
